---
title: "Powerball"
author: "David"
date: "January 21, 2016"
output: pdf_document
---
```{r, echo=FALSE}
n=as.integer(10000) ## Change the amount of tickets bought here

```
##WHY MILLIONAIRES DO NOT PLAY POWERBALL
This is a simple powerball simulater representing ONE single game of powerball where many quick pick tickets were bought.  (<http://www.powerball.com/pb_home.asp>).

```{r, echo=FALSE}
GRANDPRIZE=as.integer(60000000) #Change grand prize here
```
**Grand Prize** 

`r GRANDPRIZE`

**Winning Ticket** 
```{r, echo=FALSE}
winning.ticket=c(sort(sample(69,5)),sample(26,1))
winning.ticket
ticket=NULL
ticket=list(ticket)

powerball=rep(1:26,ceiling(n/26)+1)
#for (i in 1:n){ 
#  ticket[[i]]= c(sort(sample(69,5)),powerball[i])
#}
for (i in 1:n){ 
  ticket[[i]]= c(sort(sample(69,5)),sample(26,1))
}
sum.5w.1r=0
sum.5w=0
sum.4w.1r=0
sum.4w=0
sum.3w.1r=0
sum.3w=0
sum.2w.1r=0
sum.1w.1r=0
sum.1r=0
win.ticket=NULL
win.ticket=list(win.ticket)
for(i in 1:n){
if (sum(ticket[[i]]==winning.ticket)==6){
      sum.5w.1r=sum.5w.1r+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]==winning.ticket[1:5])==5){
      sum.5w=sum.5w+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]%in%winning.ticket[1:5])==4 && sum(ticket[[i]][6]==winning.ticket[6])==1){
      sum.4w.1r=sum.4w.1r+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]%in%winning.ticket[1:5])==4){
      sum.4w=sum.4w+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]%in%winning.ticket[1:5])==3 && sum(ticket[[i]][6]==winning.ticket[6])==1){
      sum.3w.1r=sum.3w.1r+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]%in%winning.ticket[1:5])==3){
      sum.3w=sum.3w+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]%in%winning.ticket[1:5])==2 && sum(ticket[[i]][6]==winning.ticket[6])==1){
      sum.2w.1r=sum.2w.1r+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][1:5]%in%winning.ticket[1:5])==1 && sum(ticket[[i]][6]==winning.ticket[6])==1){
      sum.1w.1r=sum.1w.1r+1
    win.ticket[[i]]=ticket[[i]]}
    else if(sum(ticket[[i]][6]==winning.ticket[6])==1){
      sum.1r=sum.1r+1
    win.ticket[[i]]=ticket[[i]]
    }
}
win.ticket=split(unlist(win.ticket), ceiling(seq_along(unlist(win.ticket))/6))
winnings=sum.5w.1r*GRANDPRIZE+sum.5w*1000000+sum.4w.1r*50000+sum.4w*100+sum.3w.1r*100+sum.3w*7+sum.2w.1r*7+sum.1w.1r*4+sum.1r*4
```
   
   The last number represents the powerball. First 5 numbers are white balls ordered from least to greatest.
   
   
 **Tickets Bought** 
   
    `r as.integer(n)`


**Winning Tickets**

For a list of all winning tickets go to appendix below.
```{r echo=FALSE, results='hide',message=FALSE}
 ## **Winning Tickets** 
for(i in 1:length(win.ticket)){
  if (sum(win.ticket[[i]]==winning.ticket)==6)
    print(paste(win.ticket[i],"5 Whites + Powerball Matched ($Grand Prize)"))
  
  else if(sum(win.ticket[[i]][1:5]==winning.ticket[1:5])==5)
    print(paste(win.ticket[i],"5 Whites Matched ($1,000,000"))
  
  else if(sum(win.ticket[[i]]%in%winning.ticket[1:5])==4 && sum(ticket[[i]][6]==winning.ticket[6])==1)
    print(paste(win.ticket[i],"4 Whites + Powerball Matched ($50,000)"))
  
  else if(sum(win.ticket[[i]][1:5]%in%winning.ticket[1:5])==4)
    print(paste(win.ticket[i],"4 Whites Matched ($100)"))
    
  else if(sum(win.ticket[[i]][1:5]%in%winning.ticket[1:5])==3 && sum(ticket[[i]][6]==winning.ticket[6])==1)
    print(paste(win.ticket[i],"3 Whites + Powerball Matched ($100)"))
  
else if(sum(win.ticket[[i]][1:5]%in%winning.ticket[1:5])==2 && sum(ticket[[i]][6]==winning.ticket[6])==1)
    print(paste(win.ticket[i],"2 Whites + Powerball Matched ($7)"))  
    
  else if(sum(win.ticket[[i]][1:5]%in%winning.ticket[1:5])==3)
    print(paste(win.ticket[i],"3 Whites Matched ($7)"))
    
  else if(sum(win.ticket[[i]][1:5]%in%winning.ticket[1:5])==1 && sum(ticket[[i]][6]==winning.ticket[6])==1)
    print(paste(win.ticket[i],"1 White + Powerball Matched ($4)"))
    
  else if(sum(win.ticket[[i]][6]==winning.ticket[6])==1)
    print(paste(win.ticket[i],"Power Ball Matched ($4)"))
}
```



**Best Ticket**
```{r, echo=FALSE}
win.ticket2=win.ticket
for(i in 1:length(win.ticket2)){
  if (sum(win.ticket2[[i]]==winning.ticket)==6)
    win.ticket2[[i]][7]=9+100
  else if(sum(win.ticket2[[i]][1:5]==winning.ticket[1:5])==5)
    win.ticket2[[i]][7]=8+100  
  else if(sum(win.ticket2[[i]]%in%winning.ticket[1:5])==4 && sum(win.ticket2[[i]][6]==winning.ticket[6])==1)
    win.ticket2[[i]][7]=7+100
  else if(sum(win.ticket2[[i]][1:5]%in%winning.ticket[1:5])==4)
    win.ticket2[[i]][7]=6+100
  else if(sum(win.ticket2[[i]][1:5]%in%winning.ticket[1:5])==3 && sum(win.ticket2[[i]][6]==winning.ticket[6])==1)
    win.ticket2[[i]][7]=5+100
  else if(sum(win.ticket2[[i]][1:5]%in%winning.ticket[1:5])==3)
    win.ticket2[[i]][7]=3+100
  else if(sum(win.ticket2[[i]][1:5]%in%winning.ticket[1:5])==2 && sum(win.ticket2[[i]][6]==winning.ticket[6])==1)
    win.ticket2[[i]][7]=4+100
  else if(sum(win.ticket2[[i]][1:5]%in%winning.ticket[1:5])==1 && sum(win.ticket2[[i]][6]==winning.ticket[6])==1)
    win.ticket2[[i]][7]=2+100
  else if(sum(win.ticket2[[i]][6]==winning.ticket[6])==1)
    win.ticket2[[i]][7]=1+100
}
po=unlist(win.ticket2)
names(po)=NULL
best.ticket=which(po[seq(7,length(win.ticket2)*7,by=7)]==max(po[seq(7,length(win.ticket2)*7,by=7)]))[1]

best.ticket=win.ticket[[best.ticket]]
la=TRUE
while(la){
       if (sum(best.ticket==winning.ticket)==6){
    print(cat(best.ticket,"     5 Whites + Powerball Matched ($Grand Prize)"))
         print(break)
       }
       if(sum(best.ticket[1:5]==winning.ticket[1:5])==5){ 
    print(cat(best.ticket,"     5 Whites Matched ($1,000,000"))
         print(break)
       }
       if(sum(best.ticket%in%winning.ticket[1:5])==4 && sum(best.ticket[6]==winning.ticket[6])==1){ 
    print(cat(best.ticket,"     4 Whites + Powerball Matched ($50,000)"))
         print(break)
       }
       if(sum(best.ticket[1:5]%in%winning.ticket[1:5])==4){ 
    print(cat(best.ticket,"     4 Whites Matched ($100)"))
         print(break)
       }
       if(sum(best.ticket[1:5]%in%winning.ticket[1:5])==3 && sum(best.ticket[6]==winning.ticket[6])==1){ 
    print(cat(best.ticket,"     3 Whites + Powerball Matched ($100)"))
         print(break)
       }
       if(sum(best.ticket[1:5]%in%winning.ticket[1:5])==2 && sum(best.ticket[6]==winning.ticket[6])==1){ 
    print(cat(best.ticket,"     2 Whites + Powerball Matched ($7)"))
         print(break)
       }
       if(sum(best.ticket[1:5]%in%winning.ticket[1:5])==3){
          cat(best.ticket,"     3 Whites Matched ($7)")
         print(break)
       }
       if(sum(best.ticket[1:5]%in%winning.ticket[1:5])==1 && sum(best.ticket[6]==winning.ticket[6])==1){ 
    print(cat(best.ticket,"     1 White + Powerball Matched ($4)"))
         print(break)
       }
       if(sum(best.ticket[6]==winning.ticket[6])==1){ 
    print(cat(best.ticket,"     Power Ball Matched ($4)"))
         print(break)
         
       }
  
}
```



```{r, echo=FALSE}
x=c(sum.5w.1r,sum.5w,sum.4w.1r,sum.4w,sum.3w.1r,sum.3w,sum.2w.1r,sum.1w.1r,sum.1r,length(win.ticket))
barplot(x,las=2,names.arg=c("5W 1R","5W","4W 1R","4W","3W 1R","3W","2W 1R","1W 1R","  1R","Total"),col=rainbow(10),main=paste("Winning Tickets"),ylab="Winning Tickets")

```
```{r, echo=FALSE}
y=c("Grand Prize","1,000,000","50,000","100","100","7","7","4","4","Total")


x=c(sum.5w.1r*GRANDPRIZE, sum.5w*1000000, sum.4w.1r*50000, sum.4w*100,sum.3w.1r*100, sum.3w*7, sum.2w.1r*7, sum.1w.1r*4, sum.1r*4,winnings)
barplot(x,las=2,names.arg=c("5W 1R","5W","4W 1R","4W","3W 1R","3W","2W 1R","1W 1R","1R","Total"),col=rainbow(10),main=paste("Ticket Winnings"),ylab="Money Won")


```
```{r, echo=FALSE}
barplot(c(2*n,winnings,abs(2*n-winnings)),col=c(44:46),names.arg=c("Spent on Tickets","Won","Lost"),ylab="Money")
```

 **Summary** 
```{r, echo=FALSE}
cat(  " You spent $", as.integer(2*n),"and won $", as.integer(winnings),"", "\n",
      as.integer(length(win.ticket)), "out of", n , "tickets were winners.","\n",
      "You lost $", (2*n)-winnings,"(if negative you won!!).","\n","\n",
      "You got 5 White + Powerball   ", sum.5w.1r, "time(s).","\n",
      "You got 5 White               ", sum.5w, "time(s).","\n",
      "You got 4 White + Powerball   ", sum.4w.1r, "time(s).","\n",
      "You got 4 White               ", sum.4w, "time(s),","\n",
      "You got 3 White + Powerball   ", sum.3w.1r, "time(s).","\n",
      "You got 3 White               ", sum.3w, "time(s),","\n",
      "You got 2 White + Powerball   ", sum.2w.1r, "time(s).","\n",
      "You got 1 White + Powerball   ", sum.1w.1r, "time(s).","\n",
      "You got Powerball             ", sum.1r, "time(s)."
      )
```
 
 **Appendix** 
 List of all winning tickets.
```{r, echo=FALSE}
win.ticket
```
**do to the way the code was written duplicate tickets are possible but highly unlikely